Tracing the Meta-Level: PyPy's Tracing JIT Compiler

### 2009-04-08 talk/icooolps2009/paper.tex
Add a sentence.
Typo.
try to streamline the beginning of the abstract a bit and kill a sentence that goes into a bit too much detail about the problem

We attempt to use the technique of Tracing JIT Compilers
(see e.g.\ \cite{andreas_gal_incremental_2006}) in the context
of the PyPy project, \ie on programs that are interpreters for some
dynamic language (including Python).  Tracing JIT compilers can greatly
speed up programs that spend most of their time in loops in which they
take similar code paths.  However, applying an unmodified tracing JIT to
a program that is itself a bytecode interpreter results in very limited
or no speedup.  One of the reasons is that their main loop, \ie the
bytecode dispatch, always executes different bytecodes and so follows
different paths.

In this paper we show how to guide tracing JIT compilers to greatly
improve the speed of bytecode interpreters. We evaluate our technique by
applying it to two PyPy interpreters: one is a small example, and the
other one is the full Python interpreter.

>>>>

We attempt to use the technique of Tracing JIT Compilers
\cite{gal_hotpathvm:effective_2006, andreas_gal_incremental_2006,
mason_chang_efficient_2007} in the context
of the PyPy project, \ie on programs that are interpreters for some
dynamic languages, including Python.  Tracing JIT compilers can greatly
speed up programs that spend most of their time in loops in which they
take similar code paths.  However, applying an unmodified tracing JIT to
a program that is itself a bytecode interpreter results in very limited
or no speedup.
In this paper we show how to guide tracing JIT compilers to greatly
improve the speed of bytecode interpreters.  One crucial point is to
unroll the bytecode dispatch loop, based on two hints provided by the
implementer of the bytecode interpreter.  We evaluate our technique by
applying it to two PyPy interpreters: one is a small example, and the
other one is the full Python interpreter.
